\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-07 paper.qmd
modifying error plots

where $n$ is the number of observations, $b$ is the number of bins, $n_h$ is the number of observations in $h^{th}$ bin, $p$ is the number of variables, $\mathbfit{x}_{hij}$ is the $j^{th}$ dimensional data of $i^{th}$ observation in $h^{th}$ hexagon.

@@

```{r}
#| label: fig-p-d-error-in-2d-two-curvy
#| fig-cap: "The $7\\text{-}D$ model error in \\gD{} layout. Color indicates square root of absolute error, dark blue indicating high error and light indicates low error. Most large errors are distributed near the edges of the clusters."
#| out-width: 30%
#| fig-pos: H

error_plot_two_curvy
```

@@

```{r}
#| echo: false
#| label: fig-bins-two-curvy
#| fig-pos: H
#| fig-cap: "Hexbin density plots of tSNE layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 80 \\text{ } (10, \\text{ }8)$, (b) $b = 130 \\text{ } (13, \\text{ }10)$, and (c) $b = 391 \\text{ } (23, \\text{ }17)$. Color indicates standardized counts, dark indicating high count and light indicates low count. At the smallest bin size, the data structure is discontinuous, suggesting that there are too many bins. Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins. XXX All look the same"
#| fig-width: 9
#| fig-height: 3

>>>>

where $n$ is the number of observations, $b$ is the number of bins, $n_h$ is the number of observations in $h^{th}$ bin, $p$ is the number of variables, $\mathbfit{x}_{hij}$ is the $j^{th}$ dimensional data of $i^{th}$ observation in $h^{th}$ hexagon. We can consider $e_{hj} = \sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2$ to be the residual for each observation.

@@

error_plot_two_curvy_hist <- 
  ggplot(error_df_two_curvy_abs) +
  geom_histogram(aes(x=sqrt_row_wise_abs_error, y=..density..)) +
  geom_density(aes(x=sqrt_row_wise_abs_error, y=..density..), colour="red") +
  xlab(expression(group("|", e[hj], "|"))) +
  ylab("") +
  theme_bw() 
```

```{r}
#| label: fig-p-d-error-in-2d-two-curvy
#| fig-cap: "The $7\\text{-}D$ model error in \\gD{} layout. Color indicates square root of absolute error, dark blue indicating high error and light indicates low error. Most large errors are distributed near the edges of the clusters."
#| out-width: 60%
#| fig-pos: H

error_plot_two_curvy + error_plot_two_curvy_hist +
  plot_layout(ncol=2)
```

@@

```{r}
#| echo: false
#| label: fig-bins-two-curvy
#| fig-pos: H
#| fig-cap: "Hexbin density plots of tSNE layout of the two C-shaped clusters data, using three different bin inputs: (a) $b = 80 \\text{ } (10, \\text{ }8)$, (b) $b = 130 \\text{ } (13, \\text{ }10)$, and (c) $b = 391 \\text{ } (23, \\text{ }17)$. Color indicates standardized counts, dark indicating high count and light indicates low count. At the smallest bin size, the data structure is discontinuous, suggesting that there are too many bins. Using the MSE of the model fit in $7\\text{-}D$ helps decide on a useful choice of number of bins."
#| fig-width: 9
#| fig-height: 3

### 2025-01-07 paper.tex
Initial writeup
modifying error plots

\includegraphics[width=0.3\textwidth,height=\textheight]{paper_files/figure-pdf/fig-hex-param-1.pdf}

@@

where \(n\) is the number of observations, \(b\) is the number of bins,
\(n_h\) is the number of observations in \(h^{th}\) bin, \(p\) is the
number of variables, \(\mathbfit{x}_{hij}\) is the \(j^{th}\)
dimensional data of \(i^{th}\) observation in \(h^{th}\) hexagon.

@@

\includegraphics[width=0.3\textwidth,height=\textheight]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}

@@

\caption{\label{fig-bins-two-curvy}Hexbin density plots of tSNE layout
of the two C-shaped clusters data, using three different bin inputs: (a)
\(b = 80 \text{ } (10, \text{ }8)\), (b)
\(b = 130 \text{ } (13, \text{ }10)\), and (c)
\(b = 391 \text{ } (23, \text{ }17)\). Color indicates standardized
counts, dark indicating high count and light indicates low count. At the
smallest bin size, the data structure is discontinuous, suggesting that
there are too many bins. Using the MSE of the model fit in
\(7\text{-}D\) helps decide on a useful choice of number of bins. XXX
All look the same}

@@

\includegraphics[width=0.6\textwidth,height=\textheight]{paper_files/figure-pdf/fig-two_non_linear_diff_shaped_close_clusters-mse-1.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-one-dens_clust-error-1.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}

@@

\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}

@@

\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}

@@

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}

@@

\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-model-error-mnist-1.pdf}

>>>>

\includegraphics[width=0.3\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-hex-param-1.pdf}

@@

where \(n\) is the number of observations, \(b\) is the number of bins,
\(n_h\) is the number of observations in \(h^{th}\) bin, \(p\) is the
number of variables, \(\mathbfit{x}_{hij}\) is the \(j^{th}\)
dimensional data of \(i^{th}\) observation in \(h^{th}\) hexagon. We can
consider
\(e_{hj} = \sum_{j = 1}^{p} (\mathbfit{x}_{hij} - C^{(p)}_{hj})^2\) to
be the residual for each observation.

@@

\includegraphics[width=0.6\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-p-d-error-in-2d-two-curvy-1.pdf}

@@

\caption{\label{fig-bins-two-curvy}Hexbin density plots of tSNE layout
of the two C-shaped clusters data, using three different bin inputs: (a)
\(b = 80 \text{ } (10, \text{ }8)\), (b)
\(b = 130 \text{ } (13, \text{ }10)\), and (c)
\(b = 391 \text{ } (23, \text{ }17)\). Color indicates standardized
counts, dark indicating high count and light indicates low count. At the
smallest bin size, the data structure is discontinuous, suggesting that
there are too many bins. Using the MSE of the model fit in
\(7\text{-}D\) helps decide on a useful choice of number of bins.}

@@

\includegraphics[width=0.6\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-two_non_linear_diff_shaped_close_clusters-mse-1.pdf}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-one-dens_clust-error-1.pdf}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pbmc-mse-1.pdf}

@@

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-model-pbmc-author-proj-1.pdf}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}

@@

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-model-pbmc-author-proj-scdeed-1.pdf}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-mnist-mse-1.pdf}

@@

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-model-error-mnist-1.pdf}
